### 2014-10-03 tex/proofs.tex
Minor revision on lem S-determ

\begin{lemma}[$\rsum^*$ preserves heap determinism]
\label{lem:S-determ}
Given a deterministic heap, $(\cfgnt{L}_0\ \cfgnt{R}_0)$, from a
state with a reference $\cfgnt{r}$ and field $\cfgnt{f}$, the
new heap, $(\cfgnt{L}^\prime\ \cfgnt{R}^\prime)$, from the 
summary machine, $(\cfgnt{L}_0\ \cfgnt{R}_0\ \cfgnt{r}\ \cfgnt{f})
\rsum^*
(\cfgnt{L}^\prime\ \cfgnt{R}^\prime\ \cfgnt{r}\ \cfgnt{f})$, is also deterministic.
\end{lemma}
\begin{proof}
Induction over the number of machine steps on the symbolic initialization machine in \figref{fig:symInit}.

\noindent\textbf{Base Case}. The machine takes one step: $(\cfgnt{L}_0\ \cfgnt{R}_0\ \cfgnt{r}\ \cfgnt{f})
\rsum
(\cfgnt{L}_1\ \cfgnt{R}_1\ \cfgnt{r}\ \cfgnt{f})$. Let $\Lambda = \mathbb{UN}(\cfgnt{L}_0, \cfgnt{R}_0, \cfgnt{r},
\cfgnt{f})$ be the set of uninitialized locations. If $\Lambda = \emptyset$, then the \texttt{Summarize (end)}
rule is active and $(\cfgnt{L}_1\ \cfgnt{R}_1) = (\cfgnt{L}_0\ \cfgnt{R}_0)$, which is deterministic by the initial conditions in the lemma.

If $\Lambda \neq \emptyset$, then the \texttt{Summarize} rule is
active, and each new constraint location pair must be considered
individually. These pairs are partitioned into the sets
$\theta_\mathit{null}$, $\theta_\mathit{new}$,
$\theta_\mathit{alias}$, and $\theta_\mathit{orig}$ by the rule. 
\begin{itemize}
\item The original heap is
  deterministic by definition, so any constraint in any member of the
  set must have some term such that
  \[\begin{array}{l}
     \forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \theta_\mathit{orig}\ \\
     \ \ \ \ ((\cfgnt{l} \neq \cfgnt{l}^\prime \vee \phi \neq \phi^\prime) \Rightarrow (\phi \wedge \phi^\prime = \cfgt{false}))
     \end{array}
  \]
  Further, any member of $\theta_\mathrm{orig}$ has a constraint
  of the form $\phi = \neg \phi_x \wedge \ldots$ while any member of $\theta_\mathit{null}$, $\theta_\mathit{new}$, and
  $\theta_\mathit{alias}$ has a constraint of the form $\phi^\prime = \phi_x \wedge
  \ldots$; thus
  \[\begin{array}{l}
     \forall (\phi\ \cfgnt{l}) \in \theta_\mathit{orig}\ (\forall (\phi^\prime\ \cfgnt{l}^\prime) \in \theta_\mathit{null} \cup \theta_\mathit{new} \cup \theta_\mathit{alias}\ (\\
     \ \ \ \ \phi \wedge \phi^\prime = \cfgt{false}))
    \end{array}
  \]
\item The only member of $\theta_\mathit{null}= \{(\phi\ \cfgnt{l}_\mathit{null})\}$ has the form $\phi = \ldots \wedge
  \cfgnt{r}_f = \cfgnt{r}_\mathit{null}$ while any member of
  $\theta_\mathit{new}$ and $\theta_\mathit{alias}$ has the form
  $\phi^\prime = \ldots \wedge \cfgnt{r}_f \neq \cfgnt{r}_\mathit{null} \wedge
  \ldots$; thus
  \[\begin{array}{l}
     \forall (\phi^\prime\ \cfgnt{l}^\prime) \in \theta_\mathit{new} \cup \theta_\mathit{alias}\ (
     \phi \wedge \phi^\prime = \cfgt{false})
    \end{array}
  \]
\item The only member of $\theta_\mathit{new} =  \{(\phi\ \cfgnt{l}_f)\}$ has a constraint of the form
  $\phi = \ldots \wedge ( \wedge_{( \cfgnt{r}_a,\ \phi_a,\ l_a) \in \rho}
  \cfgnt{r}_f \ne \cfgnt{r}_a) )$ to assert it does not alias
  anything, while any member of $\theta_\mathit{alias}$ has the form
  $\phi^\prime = \ldots \wedge \cfgnt{r}_f = \cfgnt{r}_a \wedge \ldots$ to assert it aliases some
  $\cfgnt{r}_a$ with both partitions reasoning over the same set
  of aliases $\rho$; thus
  \[\begin{array}{l}
     \forall (\phi^\prime\ \cfgnt{l}^\prime) \in \theta_\mathit{alias}\ (
     \phi \wedge \phi^\prime = \cfgt{false})
    \end{array}
  \]
\item Any member of $\theta_\mathit{alias}$ has the form
\[\ldots \wedge \cfgnt{r}_f = \cfgnt{r}_a \wedge ( \wedge_{( \cfgnt{r}^{\prime}_a\ \phi^{\prime}_a\ l^{\prime}_a)  \in \rho\ ( \cfgnt{r}^\prime_a \neq \cfgnt{r}_a) } \cfgnt{r}_f \neq \cfgnt{r}^{\prime}_a)\]
And thus, 
  \[\begin{array}{l}
     \forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \theta_\mathit{alias}\\
     \ \ \ \ ((\cfgnt{l} \neq \cfgnt{l}^\prime \vee \phi \neq \phi^\prime) \Rightarrow (\phi \wedge \phi^\prime = \cfgt{false}))
     \end{array}
  \]
\end{itemize}
As $\theta$ is mapped to a single reference $r_f$ in an already deterministic heap, the
resulting heap $(\cfgnt{L}_1\ \cfgnt{R}_1)$ is likewise deterministic.

\noindent\textbf{Inductive Step}. The machine takes $n$-steps:
\[
(\cfgnt{L}_0\ \cfgnt{R}_0\ \cfgnt{r}\ \cfgnt{f}) \rsum (\cfgnt{L}_1\ \cfgnt{R}_1\ \cfgnt{r}\ \cfgnt{f}) \rsum \ldots \rsum
(\cfgnt{L}_n\ \cfgnt{R}_n\ \cfgnt{r}\ \cfgnt{f})
\]
By the induction hypothesis, $(\cfgnt{L}_n\ \cfgnt{R}_n)$ is location
mutual exclusive. This matches the base case, in that the heap on the
left side of $\rsum$ is location mutual exclusive, and by the
same argument as in the base case,
$(\cfgnt{L}_{n+1}\ \cfgnt{R}_{n+1})$ is thus location mutual exclusive.
\end{proof}

\begin{lemma}[$\rightarrow_\mathit{FA}$ preserves heap determinism]
\label{lem:FA-determ}
Given a state, $s_\sym$, with a deterministic heap,
$(\cfgnt{L}_\sym\ \cfgnt{R}_\sym) =
\mathrm{heap}(s_\sym)$, the new heap,
$(\cfgnt{L}_\sym^\prime\ \cfgnt{R}_\sym^\prime) =
\mathrm{heap}(s_\sym^\prime)$, in any state related by the field
access rule, $s_\sym \rightarrow_\mathit{FA} s_\sym^\prime$,
is also deterministic.
\end{lemma}
\begin{proof}
Proof by definition of $\rightarrow_\mathit{FA}$ in \figref{fig:sym}.

\lemref{lem:S-determ} establishes that the heap coming out of the
$S$-relation is deterministic, so it is only needed to show that
determinism is preserved by the call to the value function,
$\mathbb{VS}$, in the rule. Let $(\cfgnt{L}\ \cfgnt{R})$ be that new
deterministic heap.

Recall from \defref{def:VS} that each constraint in each member of the value set has the form
$(\phi\wedge \phi^\prime\ \cfgnt{l})$. Choose any two distinct members of the value set, $(\phi_\alpha \wedge
\phi_\alpha^\prime\ \cfgnt{l}_\alpha^\prime)$ and $(\phi_\beta \wedge
\phi_\beta^\prime\ \cfgnt{l}_\beta^\prime)$.
\begin{itemize}
\item If $\phi_\alpha = \phi_\beta$, then by \defref{def:VS} 
\[\begin{array}{l}
\exists (\phi_\alpha\ \cfgnt{l}) \in \cfgnt{L}(\cfgnt{r})\ (\exists \cfgnt{r}^\prime \in \cfgnt{R}(\cfgnt{l},\cfgnt{f})\ (\\
\ \ \ \ (\phi_\alpha^\prime\ l_\alpha^\prime) \in \cfgnt{L}(\cfgnt{r}^\prime) \wedge (\phi_\beta^\prime\ l_\beta^\prime) \in \cfgnt{L}(\cfgnt{r}^\prime)))
\end{array}
\]
As $(\phi_\alpha^\prime\ \cfgnt{l}_\alpha^\prime)$ and $(\phi_\beta^\prime\ \cfgnt{l}_\beta^\prime)$ are distinct and connected to the same reference $\cfgnt{r}^\prime$ in a deterministic heap, $\phi_\alpha^\prime \wedge  \phi_\beta^\prime = \cfgt{false}$ by definition.
\item If $\phi_\alpha \ne \phi_\beta$, then by \defref{def:VS} 
\[\begin{array}{l}
\exists \cfgnt{l}\ ((\phi_\alpha\ \cfgnt{l}) \in \cfgnt{L}(\cfgnt{r}) \wedge \exists \cfgnt{l}^\prime \neq \cfgnt{l} \ ((\phi_\beta\ \cfgnt{l}^\prime) \in \cfgnt{L}(\cfgnt{r})))
\end{array}
\]
As $(\phi_\alpha\ \cfgnt{l})$ and $(\phi_\beta\ \cfgnt{l}^\prime)$ are distinct and connected to the same reference $\cfgnt{r}$ in a deterministic heap, $\phi_\alpha \wedge  \phi_\beta = \cfgt{false}$ by definition.
\end{itemize}
The only change to the heap after the $S$-relation is the addition of
the new reference $\cfgnt{r}^\prime = \mathrm{stack}_\cfgnt{r}()$ to
point to the value set. As the value set meets the conditions for
determinism, the new heap with $r^\prime$ and the value set,
$(\cfgnt{L}_\sym^\prime\ \cfgnt{R}_\sym^\prime) =
\mathrm{heap}(s_\sym^\prime)$, is also deterministic.
\end{proof}

>>>>

\begin{lemma}[$\rsum^*$ preserves heap determinism]
\label{lem:S-determ}
Given a deterministic heap, $(\cfgnt{L}_0\ \cfgnt{R}_0)$, from a
state with a reference $\cfgnt{r}$ and field $\cfgnt{f}$, the
new heap, $(\cfgnt{L}^\prime\ \cfgnt{R}^\prime)$, from the 
summary machine, $(\cfgnt{L}_0\ \cfgnt{R}_0\ \cfgnt{r}\ \cfgnt{f})
\rsum^*
(\cfgnt{L}^\prime\ \cfgnt{R}^\prime\ \cfgnt{r}\ \cfgnt{f})$, is also deterministic.
\end{lemma}
\begin{proof}
Induction over the number of steps in $\rsum^*$ in \figref{fig:symInit}.

\noindent\textbf{Base Case}. The relation makes one step: $(\cfgnt{L}_0\ \cfgnt{R}_0\ \cfgnt{r}\ \cfgnt{f})
\rsum
(\cfgnt{L}_1\ \cfgnt{R}_1\ \cfgnt{r}\ \cfgnt{f})$. Let $\Lambda = \mathbb{UN}(\cfgnt{L}_0, \cfgnt{R}_0, \cfgnt{r},
\cfgnt{f})$ be the set of uninitialized locations. If $\Lambda = \emptyset$, then the \textrm{S{\footnotesize UMMARIZE-END}}
rule is active and $(\cfgnt{L}_1\ \cfgnt{R}_1) = (\cfgnt{L}_0\ \cfgnt{R}_0)$, which is deterministic by the initial conditions in the lemma.

If $\Lambda \neq \emptyset$, then the \textrm{S{\footnotesize UMMARIZE}} rule is
active, and each new constraint location pair must be considered
individually. These pairs are partitioned into the sets
$\theta_\mathit{null}$, $\theta_\mathit{new}$,
$\theta_\mathit{alias}$, and $\theta_\mathit{orig}$ by the rule. 
\begin{itemize}
\item The original heap is
  deterministic by definition, so any constraint in any member of the
  set must have some term such that
  \[\begin{array}{l}
     \forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \theta_\mathit{orig}\ \\
     \ \ \ \ ((\cfgnt{l} \neq \cfgnt{l}^\prime \vee \phi \neq \phi^\prime) \Rightarrow (\phi \wedge \phi^\prime = \cfgt{false}))
     \end{array}
  \]
  Further, any member of $\theta_\mathrm{orig}$ has a constraint
  of the form $\phi = \neg \phi_x \wedge \ldots$ while any member of $\theta_\mathit{null}$, $\theta_\mathit{new}$, and
  $\theta_\mathit{alias}$ has a constraint of the form $\phi^\prime = \phi_x \wedge
  \ldots$; thus
  \[\begin{array}{l}
     \forall (\phi\ \cfgnt{l}) \in \theta_\mathit{orig}\ (\forall (\phi^\prime\ \cfgnt{l}^\prime) \in \theta_\mathit{null} \cup \theta_\mathit{new} \cup \theta_\mathit{alias}\ (\\
     \ \ \ \ \phi \wedge \phi^\prime = \cfgt{false}))
    \end{array}
  \]
\item The only member of $\theta_\mathit{null}= \{(\phi\ \cfgnt{l}_\mathit{null})\}$ has the form $\phi = \ldots \wedge
  \cfgnt{r}_f = \cfgnt{r}_\mathit{null}$ while any member of
  $\theta_\mathit{new}$ and $\theta_\mathit{alias}$ has the form
  $\phi^\prime = \ldots \wedge \cfgnt{r}_f \neq \cfgnt{r}_\mathit{null} \wedge
  \ldots$; thus
  \[\begin{array}{l}
     \forall (\phi^\prime\ \cfgnt{l}^\prime) \in \theta_\mathit{new} \cup \theta_\mathit{alias}\ (
     \phi \wedge \phi^\prime = \cfgt{false})
    \end{array}
  \]
\item The only member of $\theta_\mathit{new} =  \{(\phi\ \cfgnt{l}_f)\}$ has a constraint of the form
  $\phi = \ldots \wedge ( \wedge_{( \cfgnt{r}_a,\ \phi_a,\ l_a) \in \rho}
  \cfgnt{r}_f \ne \cfgnt{r}_a) )$ to assert it does not alias
  anything, while any member of $\theta_\mathit{alias}$ has the form
  $\phi^\prime = \ldots \wedge \cfgnt{r}_f = \cfgnt{r}_a \wedge \ldots$ to assert it aliases some
  $\cfgnt{r}_a$ with both partitions reasoning over the same set
  of aliases $\rho$; thus
  \[\begin{array}{l}
     \forall (\phi^\prime\ \cfgnt{l}^\prime) \in \theta_\mathit{alias}\ (
     \phi \wedge \phi^\prime = \cfgt{false})
    \end{array}
  \]
\item Any member of $\theta_\mathit{alias}$ has the form
\[\ldots \wedge \cfgnt{r}_f = \cfgnt{r}_a \wedge ( \wedge_{( \cfgnt{r}^{\prime}_a\ \phi^{\prime}_a\ l^{\prime}_a)  \in \rho\ ( \cfgnt{r}^\prime_a \neq \cfgnt{r}_a) } \cfgnt{r}_f \neq \cfgnt{r}^{\prime}_a)\]
And thus, 
  \[\begin{array}{l}
     \forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \theta_\mathit{alias}\\
     \ \ \ \ ((\cfgnt{l} \neq \cfgnt{l}^\prime \vee \phi \neq \phi^\prime) \Rightarrow (\phi \wedge \phi^\prime = \cfgt{false}))
     \end{array}
  \]
\end{itemize}
As $\theta$ is mapped to a single reference $r_f = \mathit{init}_r()$ in an already deterministic heap, the
resulting heap $(\cfgnt{L}_1\ \cfgnt{R}_1)$ is likewise deterministic.

\noindent\textbf{Inductive Step}. The machine takes $n$-steps:
\[
(\cfgnt{L}_0\ \cfgnt{R}_0\ \cfgnt{r}\ \cfgnt{f}) \rsum (\cfgnt{L}_1\ \cfgnt{R}_1\ \cfgnt{r}\ \cfgnt{f}) \rsum \ldots \rsum
(\cfgnt{L}_n\ \cfgnt{R}_n\ \cfgnt{r}\ \cfgnt{f})
\]
By the induction hypothesis, $(\cfgnt{L}_n\ \cfgnt{R}_n)$ is deterministic. This matches the base case, in that the heap on the
left side of $\rsum$ is deterministic, and by the
same argument as in the base case,
$(\cfgnt{L}_{n+1}\ \cfgnt{R}_{n+1})$ is thus deterministic.
\end{proof}

\begin{lemma}[$\rightarrow_\mathit{FA}$ preserves heap determinism]
\label{lem:FA-determ}
Given a state, $s$, with a deterministic heap,
$(\cfgnt{L}\ \cfgnt{R}) =
\mathrm{heap}(s)$, the new heap,
$(\cfgnt{L}^\prime\ \cfgnt{R}^\prime) =
\mathrm{heap}(s^\prime)$, in any state related by the field
access rule, $s \rightarrow_\mathit{FA} s^\prime$,
is also deterministic.
\end{lemma}
\begin{proof}
Proof by definition of $\rightarrow_\mathit{FA}$ in \figref{fig:sym}.

\lemref{lem:S-determ} establishes that the heap in the state on the right side of
$\rsum^*$ is deterministic if the heap in the state on the left side
is deterministic, so it is only needed to show that determinism is
preserved by the call to the value function, $\mathbb{VS}$, in the
rule. Let $(\cfgnt{L}\ \cfgnt{R})$ be the new deterministic heap related by $\rsum^*$.

Recall from \defref{def:VS} that each constraint in each member of the value set has the form
$(\phi\wedge \phi^\prime\ \cfgnt{l})$. Choose any two distinct members of the value set, $(\phi_\alpha \wedge
\phi_\alpha^\prime\ \cfgnt{l}_\alpha^\prime)$ and $(\phi_\beta \wedge
\phi_\beta^\prime\ \cfgnt{l}_\beta^\prime)$.
\begin{itemize}
\item If $\phi_\alpha = \phi_\beta$, then by \defref{def:VS} 
\[\begin{array}{l}
\exists (\phi_\alpha\ \cfgnt{l}) \in \cfgnt{L}(\cfgnt{r})\ (\exists \cfgnt{r}^\prime \in \cfgnt{R}(\cfgnt{l},\cfgnt{f})\ (\\
\ \ \ \ (\phi_\alpha^\prime\ l_\alpha^\prime) \in \cfgnt{L}(\cfgnt{r}^\prime) \wedge (\phi_\beta^\prime\ l_\beta^\prime) \in \cfgnt{L}(\cfgnt{r}^\prime)))
\end{array}
\]
As $(\phi_\alpha^\prime\ \cfgnt{l}_\alpha^\prime)$ and $(\phi_\beta^\prime\ \cfgnt{l}_\beta^\prime)$ are distinct and connected to the same reference $\cfgnt{r}^\prime$ in a deterministic heap, $\phi_\alpha^\prime \wedge  \phi_\beta^\prime = \cfgt{false}$ by definition.
\item If $\phi_\alpha \ne \phi_\beta$, then by \defref{def:VS} 
\[\begin{array}{l}
\exists \cfgnt{l}\ ((\phi_\alpha\ \cfgnt{l}) \in \cfgnt{L}(\cfgnt{r}) \wedge \exists \cfgnt{l}^\prime \neq \cfgnt{l} \ ((\phi_\beta\ \cfgnt{l}^\prime) \in \cfgnt{L}(\cfgnt{r})))
\end{array}
\]
As $(\phi_\alpha\ \cfgnt{l})$ and $(\phi_\beta\ \cfgnt{l}^\prime)$ are distinct and connected to the same reference $\cfgnt{r}$ in a deterministic heap, $\phi_\alpha \wedge  \phi_\beta = \cfgt{false}$ by definition.
\end{itemize}
The only change to the heap after the $S$-relation is the addition of
the new reference $\cfgnt{r}^\prime = \mathrm{stack}_\cfgnt{r}()$ to
point to the value set. As the value set meets the conditions for
determinism, the new heap with $r^\prime$ and the value set,
$(\cfgnt{L}^\prime\ \cfgnt{R}^\prime) =
\mathrm{heap}(s^\prime)$, is also deterministic.
\end{proof}
